---
title: 'Mileage vs Transmission type : Regression Models(Coursera)'
author: "Debjyoti Paul"
date: "Sunday, December 21, 2014"
output: pdf_document
keep_md: yes
---
Executive Summary
--------------------------------------------------------
In this project, the dependency of mileage on Transmission (automatic or manual) in an automobile is analyzed using **mtcars** data <https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html>. By fitting a regression model, inference is drawn on the influence on mileage due to transmission mode.

Analysis
--------------------------------------------------------
**mtcars** data frame is loaded and  the data in **am** variable is stored as a logical **autm** variable.

```{r echo=TRUE, eval=TRUE}
data( mtcars )
mtcars$autm <- mtcars$am == 0
attach( mtcars )
```

Some exploratory data analysis on the distribution of the automatic variable is depicted in (Appendix).

Initially the data is analzed with the Ordinary Least Squares (OLS) model.

```{r echo=TRUE, eval=TRUE}
ls <- lm( mpg ~ autm, data=mtcars )
summary( ls )
```

The estimation clearly indicates a strong negative influence on mileage due to automatic transmission.

The model is estimated using **Weighted Least Squares** (WLS), using the weight (**wt**) varible to facilitate better fitting. 

```{r echo=TRUE, eval=TRUE}

wls = lm (mpg ~ autm, data=mtcars,weights=(1/mtcars$wt^2))

summary(wls)
```

The summary depicts a stronger negative influence and the coeffecients are more significant than the that with Ordinary Least Square. The residuals for both estimations is plotted in (Appendix).



Appendix: Figures
--------------------------------------------------------
```{r echo=TRUE, eval=TRUE, fig.width=10, fig.height=3}
par( mfrow=c(1,2) )
hist( mpg[which(autm == TRUE)] , main="Distribution of mileage for automatic transmission",xlab="mileage(mpg)")
hist( mpg[which(autm == FALSE)],main="Distribution of mileage for manual transmission",xlab="mileage(mpg)" )
```
There is a strong correlation between mode of transmission and the weight of the automible. This dependence can be exploited as wieghts to yeild better results.

--------------------------------------------------------
```{r echo=FALSE, eval=TRUE, fig.width=10, fig.height=3}
plot( mtcars$autm, mtcars$wt, xlab = 'Transmission mode', ylab = 'Weight')
```

The residuals of the Ordinary Least Square and Weighted Least Sqaure estimations.

```{r echo=TRUE, eval=TRUE, fig.width=10, fig.height=3}
par( mfrow=c(1,2) )
hist( ls$residuals , main="Distribution of residuals of ordinary least square",xlab="residuals" )
hist(wls$residuals , main="Distribution of residuals of weighted least square",xlab="residuals")

```